*Prochlorococcus marinus* responses to light and oxygen

### 2024-02-23 Code/Manuscript_O2.Rmd
Edits to Results & Discussion

We then analyzed growth rates in terms of cumulative diel Photosynthetically Usable Radiation (PUR). Photosynthetic organisms absorb light energy within the Photosynthetically Active Radiation (PAR) range, 350 to 700 nm, for photosynthesis [@morelAvailableUsableStored1978]. PUR represents the fraction of PAR that is absorbed by the pigments of a given photosynthetic organism [@morelAvailableUsableStored1978], taking into account the specific spectral wavebands these pigments absorb.  *P. marinus* light-harvesting complexes, [@chisholmProchlorococcusMarinusNov1992;
@goerickePigmentsProchlorococcusMarinus1992], show an absorption maxima
of 442 nm for divinyl chlorophyll a and 478 nm for divinyl chlorophyll b
[@goerickePigmentsProchlorococcusMarinus1992] allowing *P. marinus* to
efficiently harvest blue light in the 400 nm
to 500 nm range [@morelAvailableUsableStored1978] prevailing in deep ocean
habitats, where only blue spectral wavelengths prevail [@holtropVibrationalModesWater2021].  In *P. marinus* small cell diameters, from 0.5 to 0.7 µm [@chisholmProchlorococcusMarinusNov1992], and simple cell structures, minimize the complication of pigment package effect or intracellular self-shading [@morelProchlorococcusSynechococcusComparative1993]. Given the different spectral light regimes typical of the niches of different ecotypes, expressing growth rates in terms of cumulative diel absorbed PUR might simplify different photoperiods, spectral bands, and PAR levels into a common parameter, making growth response comparisons across strains and different oxygen levels more accessible.  We also aimed to detect whether growth responses are driven simply by cumulative diel absorbed PUR, or whether specific photoperiods, spectral bands or PAR levels have independent, albeit interacting, effects on growth.

@@

Actinic absorbed PUR levels (µmol photons m^-2^ s^-1^) were calculated from actinic PAR (µmol photons m^-2^ s^-1^) levels using the equation \@ref(eq:PUR) from [@morelAvailableUsableStored1978],

@@

The applied photoperiods were delivered using the sine circadian light function of the PSI Multicultivator to simulate light exposure approximating sun rise through to sunset. The area under the sinusoidal curves is equivalent to the area of a triangular photoregime of equivalent photoperiod (Campbell, unpub), therefore the equation to determine the cumulative diel absorbed PUR (µmol photons m^-2^ d^-1^) is one half of the base (photoperiod) multiplied by the height (PUR) (equation \@ref(eq:SumPURDay)),

\begin{equation}
  Cumulative diel absorbed PUR = \frac{PUR × 3600 × Photoperiod}{2}
  (\#eq:SumPURDay)
\end{equation}

@@

Figures \@ref(fig:MED4OverlayPlots), \@ref(fig:SS120OverlayPlots) and
\@ref(fig:MIT9313OverlayPlots) provide visual representations of PUR, the black solid line and shaded area, in relation to the imposed PAR, the dotted line, under each imposed spectral wavebands for *P. marinus* MED4, SS120 and MIT9313, respectively. Figure (\@ref(fig:PurParPlot)) shows the relationship between calculated absorbed PUR versus imposed PAR for each *P. marinus* and each spectral waveband.

@@

PurParPlot_cap <- glue("**Absorbed peak Photosynthetically Usable Radiation (PUR) (µmol photons m^-2^ s^-1^) vs. peak Photosynthetically Active Radiation (PAR) (µmol photons m^-2^ s^-1^).**",
                       "The correlation between PAR, plotted on the x-axis and absorbed PUR, plotted on the y-axis, are colored for each growth spectral waveband; 450 nm (blue circles), 660 nm (red circles) and white LED (black circles).",
                       "The grey dashed line represents a hypothetical one to one correlation.",
                       "**A.** is *Prochlorococcus marinus*  MED4.",
                       "**B.** is *Prochlorococcus marinus*  SS120.",
                       "**C.** is *Prochlorococcus marinus*  MIT9313.",

@@

Under 250 µM O~2~ growth rates of *P. marinus* SS120, Clade LLIII, increased with longer photoperiods, under 30 µmol photons m^-2^ s^-1^ peak PAR and (Figure \@ref(fig:SS120PAR)). No growth was observed under any blue light photoperiods when exposed to peak PAR of 90 µmol photons m^-2^ s^-1^ or greater. Growth rate increased with increasing photoperiods for white and red light under peak PAR of 90 µmol photons m^-2^ s^-1^.  Growth rate decreased with longer photoperiods and showed growth inhibition at 12 and 16 h photoperiods under PAR of 180 µmol photons m^-2^ s^-1^  white, red or blue LED light.  The µ~max~ was `r PURPARFits %>% filter(Strain == "SS120", WL == "WW", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^  achieved under 90 µmol photons m^-2^ s^-1^ white LED light and 16 h photoperiod.

Under 25 µM O~2~ and PAR of 30 µmol photons m^-2^ s^-1^growth trends were similar to 250 µM O~2~. SS120 maintaine showed no
growth under a 4 h photoperiod for red spectral waveband, however under blue light, SS120 was able to grow under a 4 h photoperiod (Figure
\@ref(fig:SS120PAR)). In contrast to the growth trends of the 250 µM O~2~ and PAR of 90 µmol photons m^-2^ s^-1^ experiments, SS120 grew under 4 and 8 h blue light and 16 h red light photoperiods. The growth rate decreased with increasing photoperiod under PAR of 90 µmol photons
m^-2^ s^-1^ white LED light treatments. Blue light treatments under PAR of 180 µmol photons m^-2^ s^-1^ showed growth only under an 8 h
photoperiod. The µ~max~ was `r PURPARFits %>% filter(Strain == "SS120", WL == "450", O2_uM == "25", Photoperiod == 8) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ (Table. \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons
m^-2^ s^-1^ blue light and 8 h photoperiod. The 25 µM O~2~, less than 16 h photoperiod and 180 µmol photons m^-2^ s^-1^  under white LED light experiments were not performed due to time constraints.

SS120 did not reproducibly grow when sparged to the lowest O~2~ of 2.5 µM (Figure \@ref(fig:SS120PAR)). 2.5 µM O~2~ growth experiments were not conducted for 4 and 16 h photoperiods under PAR of 180 µmol photons m^-2^ s^-1^, as no growth occurred when SS120 was exposed to 8 and 12 h photoperiods. Red light 16 h photoperiod experiments were not performed due to time constraints.

@@

The GAM model in Figure \@ref(fig:SS120GAM) summarizes growth
responses of SS120 to red (A) or blue (B) peak PAR and photoperiod,
across the 2 imposed oxygen concentrations. Under 250 µM O~2~,
Figure \@ref(fig:SS120GAM)A showed highest growth
rates below PAR of 50 µmol photons m^-2^ s^-1^ and
photoperiods between 8 and 12 h, indicated by the contour line labeled
0.29 d^-1^ (representing the 90^th^ percentile of achieved growth rate). The disjunct regions of the GAM plot results from variable growth success of SS120 under 250 µM O~2~.  Growth rate patterns under red light were similar, although somewhat faster.
In contrast, under 25 µM O~2~ and a photoperiod of 8 h SS120 exploited all blue peak PAR levels, achieving faster growth rates at a higher PAR of \~100 µmol photons
m^-2^ s^-1^, indicated by the contour line labeled 0.5 d^-1^
(representing the 90^th^ percentile of achieved growth rate), out pacing
the 90^th^ percentile fastest growth rates under 250 µM O~2~. Under red light and 25 µM O~2~ light SS120 grew across most conditions of peak PAR and photoperiod, achieving fastest growth under long photoperiods and peak PAR below 50 \~100 µmol photons
m^-2^ s^-1^.  Thus, the designation of SS120 as a LL strain is dependent upon the [O~2]. SS120 did not, however, grow reliably under tested conditions at 2.5 µM O~2~.

The LLIII clade, *P. marinus* SS120, grew under photoperiods longer than
4 h and showed increasing growth rate with increasing photoperiods.
Therefore, SS120 can potentially exploit the longer summer photoperiods
in temperate zones but again cannot sustain growth in winter through its
requirement for more than 4 h of light per day, even if water temperature warms into the Clade LLIII tolerance range.

SS120, grew under a PAR of 30 µmol photons m^-2^ s^-1^, and under some conditions at PAR of 90 µmol photons m^-2^ s^-1^, but growth was generally inhibited under PAR of 180 µmol photons
m^-2^ s^-1^. These findings align with the results reported by
@mooreComparativePhysiologySynechococcus1995 and are expected for a low
light clade. Given that SS120 can tolerate longer photoperiods and low
PAR levels, the deep ocean temperate zones are a viable habitat but not
during temperate winter due to shorter photoperiods.

SS120 exhibited different growth responses to spectral wavebands
under varying O~2~ conditions, as such, this section will address the
growth response under 250 µM O~2~ experiments, while the interactions of
spectral wavebands under all O~2~ treatments will be discussed in the
next section. SS120, grew under all spectral wavebands at a PAR of 30
µmol photons m^-2^ s^-1^. Growth was completely inhibited under blue
light at PAR levels of 90 and 180 µmol photons m^-2^ s^-1^, but growth
occurred at PAR of 90 µmol photons m^-2^ s^-1^ under all other spectral
wavebands. Generally, growth was strongly inhibited under PAR of 180
µmol photons m^-2^ s^-1^ for all spectral wavebands. The ability of SS120 to grow under low blue light and longer photoperiods, makes the deep ocean waters of the temperate zone a viable potential habitat. Under 250 µM O~2~ SS120 is constrained to deeper ocean waters through its intolerance of higher blue PAR levels.

@@

#### High oxygen environment {.unnumbered}

The growth rates of *P. marinus* MIT9313, Clade LLIV increased with longer
photoperiods under low 30 µmol photons m^-2^ s^-1^ peak PAR, under 250 µM O~2~ (Figure
\@ref(fig:MIT9313PAR)).  Under intermediate 90 µmol photons m^-2^ s^-1^ peak PAR growth rates decreased with increasing blue light photoperiod. Blue light did not induce growth at 180 µmol photons m^-2^ s^-1^ peak PAR, while MIT9313 showed only marginal growth under white LED and red light under the 8 h photoperiod, consistent with [@moorePhotophysiologyMarineCyanobacterium1999]. The µ~max~ was `r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under 30 µmol photons m^-2^ s^-1^ blue light and 16 h photoperiod.

#### Intermediate oxygen environment {.unnumbered}

For MIT9313 maintained at 25 µM O~2~ growth rate increased with increasing photoperiods for all spectral wavebands tested (Figure \@ref(fig:MIT9313PAR)). Under peak PAR of 90 µmol photons m^-2^ s^-1^ growth rate increased with longer photoperiods, with the fastest overall growth rate for MIT9313 
`r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "25", Photoperiod == 16, Par_ue == 90) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under 16 h blue light photoperiod. In contrast to the 250 µM O~2~ growth experiments, MIT9313 grew when exposed to peak PAR of 180 µmol photons m^-2^ s^-1^ and blue light under all photoperiods except 16 h; additionally,
white LED and red light treatments induced growth across all tested photoperiods under 25 µM O~2~. The 25 µM O~2~, 4 h photoperiod experiments under white LED light and were not performed due to time constraints.

#### Low oxygen environment {.unnumbered}

MIT9313 grew under 2.5 µM O~2~  particularly under blue LED light, albeit
generally slower than under the parallel experiments at 25 µM O~2~  (Figure
\@ref(fig:MIT9313PAR)). Growth estimates did show scatter among replicates, suggesting  2.5 µM O~2~  is near the tolerance limit for growth of MIT9313. Growth rates increased with longer photoperiods under blue
light treatments under peak PAR of 90 µmol photons m^-2^ s^-1^ . Growth  for MIT9313 under PAR of 180 µmol photons m^-2^ s^-1^ and blue light treatment decreased with increasing photoperiods with full growth inhibition under a 16 h photoperiod. The red light peak PAR of 180 µmol photons m^-2^ s^-1^ showed similar growth rates for 8 and 12 h photoperiods. The µ~max~ was `r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "2.5", Photoperiod == 4, Par_ue == 180) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under 4 h blue light photoperiod and PAR of 180 µmol photons m^-2^ s^-1^.  XXXXTHIS DOES NOT LOOK correct; 12 h, 90 uE photons blue light looks higher. The 2.5 µM O~2~ white LED treatments under 4, 8 and 16 h photoperiods and red light under 4 and 16 h photoperiods were not performed due to time constraints.

@@

The LLIV clade, *P. marinus* MIT9313, exploited all tested photoperiods,
and showed increasing growth rate with increasing photoperiods. Hence,
MIT9313 can potentially inhabit temperate zones with shorter winter and
longer summer photoperiods, and exploit habitats at depth subject to
photoperiod attenuation.

MIT9313, sometimes grew under all PAR levels but generally exhibited slower growth rate under PAR of 180 µmol photons m^-2^ s^-1^ and then only under some conditions. MIT9313 is well suited to inhabit the deep ocean in temperate zones with wide tolerance for photoperiod yet can also tolerate depths closer to the surface through its partial tolerance of higher PAR levels.

MIT9313, under 250 µM O~2~ grew under all spectral wavebands when exposed to PAR of 30 µmol photons m^-2^ s^-1^ and 90 µmol photons m^-2^ s^-1^ and was even able to grow under PAR of 180 µmol photons
m^-2^ s^-1^ for all spectral wavebands except blue light. MIT9313 can
thus exploit blue light in deeper waters of temperate zones and can even
tolerate moderate depths to exploit higher blue light levels, as long as
the photoperiod is shorter than 12 h.

@@

Cumulative diel absorbed PUR can potentially collapse photoperiod, PAR and spectral wavebands to a common metric of photosynthetically active light absorbed per day. Cumulative daily absorbed PUR dose (µmol photons m^-2^ d^-1^) was calculated from the imposed absorbed PUR (µmol photons m^-2^ s^-1^) and photoperiod (h). We plotted growth rates vs. cumulative diel absorbed PUR to determine whether growth is a simple response to diel PUR, across imposed photoperiods and spectral wavebands, or whether photoperiod or spectral wavebands have specific or interactive influences on growth.

Due to the absorption of *P. marinus* pigments in the blue spectral waveband range, the maximum imposed cumulative diel absorbed PUR under blue light is almost 3 times that of white LED light, and about 6 times the red light treatment, despite being derived from the same imposed photoperiods and peak PAR regimes. As such, only blue light experiments extend beyond cumulative diel PUR of ~ 2 x 10^6^ µmol photons m^-2^ d^-1^.

@@

The representative of HLI clade, *P. marinus* MED4, showed no growth under any 4 h photoperiod treatments, even when a 4 h photoperiod delivered cumulative diel PUR equivalent to other photoperiod treatments. In parallel MED4 showed no growth under 2.5 µM O~2~, no matter the level of diel cumulative PUR. 
In contrast, under 250 or 25  µM O~2~, and under any photoperiod greater than 4 h, MED4 growth was well described by a saturating response (REFERENCE for EQUATION) of growth to increasing cumulative diel PUR, with saturation of growth rate achieved around  1.0 x 10^6^ µmol m^-2^d^-1^ (Figure \@ref(fig:AllStrainPurFitsPlots)A), and no evidence of inhibition of growth at any achieved cumulative diel PUR.   XXXXDISTINCT PHOTOPERIOD FITSXXXXX

@@

The representative of the LLIII clade, *P. marinus* SS120 showed almost no growth under 2.5 µM O~2~ experiments. A majority of the 4 h photoperiod treatments did not grow under 250 µM O~2~ experiments, even when a 4 h photoperiod delivered cumulative diel PUR equivalent to other photoperiod treatments. SS120 did not grow when exposed to more than ~1.0 x 10^6^ µmol photons m^-2^ d^-1^ of cumulative diel PUR under any spectral waveband or photoperiod combination, under 250 µM O~2~ experiments(Figure \@ref(fig:AllStrainPurFitsPlots)B).

Under both 25 or 250 µM O~2~ experiments, SS120 growth plateaued by about 5.0 x 10^5^ µmol photons m^-2^ d^-1^ diel PUR, with some scatter among photoperiod and spectral waveband regimes. The onset of growth inhibition extended to higher cumulative diel PUR for cultures under 25 µM O~2~, showing that SS120 is partially protected from photoinhibition of growth by 25 µM O~2~. XXXXDISTINCT PHOTOPERIOD FITSXXXXX

@@

The LLIV clade representative, *P. marinus* MIT9313, showed growth rising to a plateau by about 5 x 10^5^ µmol photons m^-2^ d^-1^ of cumulative diel PUR, with higher growth rates over a narrower plateau under 25 and 250 µM O~2~, compared to a wider, flatter response to cumulative diel PUR under 2.5 µM O~2~. Above about 1.0 x 10^6^ µmol photons m^-2^ d^-1^ of cumulative PUR under 250 µM O~2~, MIT9313 showed full inhibition of growth, across photoperiods, and spectral wavebands.  In contrast, under 25 µM O~2~, MIT9313 showed a greatly extended exploitation of higher cumulative diel PUR, with full growth inhibition only above about 3.5 x 10^6^ µmol photons m^-2^ d^-1^.  Similarly, under 2.5 µM O~2~, MIT9313 grew more slowly, but only showed full growth inhibition above about 3.5 x 10^6^ µmol photons m^-2^ d^-1^ cumulative diel PUR (Figure \@ref(fig:AllStrainPurFitsPlots)C).  XXXXDISTINCT PHOTOPERIOD FITSXXXXX

XXXXXCHECK WHEN MIREILLE ADDS Colour specific fitsXXX This inhibition initially suggested that blue wavebands triggered photoinhibition at significantly lower levels of Photosynthetically Active Radiation (PAR) compared to other spectral wavebands. However, considering the equivalent cumulative diel absorbed PUR of the blue waveband, all treatments including white LED, red, and blue waveband exhibited approximately the same growth saturation. XXXX

Our data support enhanced growth under conditions of low cumulative diel  PUR and 660 nm (red) spectral bandwidth, as supported by [@murphyPhotoinactivationPhotosystemII2017] who found a lower cost of growth, due to decreased photoinactivation of PSII under red, compared to blue wavebands.

>>>>

We then analyzed growth rates in terms of cumulative diel Photosynthetically Usable Radiation (PUR). Photosynthetic organisms absorb light energy within the Photosynthetically Active Radiation (PAR) range, 350 to 700 nm, for photosynthesis [@morelAvailableUsableStored1978]. PUR represents the fraction of PAR that is absorbed by the pigments of a given photosynthetic organism [@morelAvailableUsableStored1978], taking into account the specific spectral wavebands these pigments absorb.  *P. marinus* light-harvesting complexes, [@chisholmProchlorococcusMarinusNov1992;
@goerickePigmentsProchlorococcusMarinus1992], show an absorption maxima
of 442 nm for divinyl chlorophyll a and 478 nm for divinyl chlorophyll b
[@goerickePigmentsProchlorococcusMarinus1992] allowing *P. marinus* to
efficiently harvest blue light in the 400 nm
to 500 nm range [@morelAvailableUsableStored1978] prevailing in deep ocean
habitats, where only blue spectral wavelengths prevail [@holtropVibrationalModesWater2021].  In *P. marinus* small cell diameters, from 0.5 to 0.7 µm [@chisholmProchlorococcusMarinusNov1992], and simple cell structures, minimize the complication of pigment package effect or intracellular self-shading [@morelProchlorococcusSynechococcusComparative1993]. Given the different spectral light regimes typical of the niches of different ecotypes, expressing growth rates in terms of cumulative diel PUR might simplify different photoperiods, spectral bands, and PAR levels into a common parameter, making growth response comparisons across strains and different oxygen levels more accessible.  We also aimed to detect whether growth responses are driven simply by cumulative diel PUR, or whether specific photoperiods, spectral bands or PAR levels have independent, albeit interacting, effects on growth.

@@

Actinic PUR levels (µmol photons m^-2^ s^-1^) were calculated from actinic PAR (µmol photons m^-2^ s^-1^) levels using the equation \@ref(eq:PUR) from [@morelAvailableUsableStored1978],

@@

The applied photoperiods were delivered using the sine circadian light function of the PSI Multicultivator to simulate light exposure approximating sun rise through to sunset. The area under the sinusoidal curves is equivalent to the area of a triangular photoregime of equivalent photoperiod (Campbell, unpub), therefore the equation to determine the cumulative diel PUR (µmol photons m^-2^ d^-1^) is one half of the base (photoperiod) multiplied by the height (PUR) (equation \@ref(eq:SumPURDay)),

\begin{equation}
  Cumulative diel PUR = \frac{PUR × 3600 × Photoperiod}{2}
  (\#eq:SumPURDay)
\end{equation}

@@

Figures \@ref(fig:MED4OverlayPlots), \@ref(fig:SS120OverlayPlots) and
\@ref(fig:MIT9313OverlayPlots) provide visual representations of PUR, the black solid line and shaded area, in relation to the imposed PAR, the dotted line, under each imposed spectral wavebands for *P. marinus* MED4, SS120 and MIT9313, respectively. Figure (\@ref(fig:PurParPlot)) shows the relationship between calculated PUR versus imposed PAR for each *P. marinus* and each spectral waveband.

@@

PurParPlot_cap <- glue("**Absorbed peak Photosynthetically Usable Radiation (PUR) (µmol photons m^-2^ s^-1^) vs. peak Photosynthetically Active Radiation (PAR) (µmol photons m^-2^ s^-1^).**",
                       "The correlation between PAR, plotted on the x-axis and PUR, plotted on the y-axis, are colored for each growth spectral waveband; 450 nm (blue circles), 660 nm (red circles) and white LED (black circles).",
                       "The grey dashed line represents a hypothetical one to one correlation.",
                       "**A.** is *Prochlorococcus marinus*  MED4.",
                       "**B.** is *Prochlorococcus marinus*  SS120.",
                       "**C.** is *Prochlorococcus marinus*  MIT9313.",

@@

Under 250 µM O~2~ growth rates of *P. marinus* SS120, Clade LLIII, increased with longer photoperiods, under 30 µmol photons m^-2^ s^-1^ peak PAR and (Figure \@ref(fig:SS120PAR)). No growth was observed under any blue light photoperiods when exposed to peak PAR of 90 µmol photons m^-2^ s^-1^ or greater. Growth rate, however increased with increasing photoperiods for white and red light under peak PAR of 90 µmol photons m^-2^ s^-1^.  Growth rate decreased with longer photoperiods and showed growth inhibition at 12 and 16 h photoperiods under PAR of 180 µmol photons m^-2^ s^-1^  white, red or blue LED light.  The µ~max~ was `r PURPARFits %>% filter(Strain == "SS120", WL == "WW", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^  achieved under 90 µmol photons m^-2^ s^-1^ white LED light and 16 h photoperiod.

Under 25 µM O~2~ and PAR of 30 µmol photons m^-2^ s^-1^ growth trends were similar to 250 µM O~2~. SS120 showed no
growth under a 4 h photoperiod for red spectral waveband, however under blue light, SS120 was able to grow under a 4 h photoperiod (Figure
\@ref(fig:SS120PAR)). In contrast to the growth trends of the 250 µM O~2~ and PAR of 90 µmol photons m^-2^ s^-1^ experiments, SS120 grew under 4 and 8 h blue light and 16 h red light photoperiods. The growth rate decreased with increasing photoperiod under PAR of 90 µmol photons m^-2^ s^-1^ white LED light treatments. Blue light treatments under PAR of 180 µmol photons m^-2^ s^-1^ showed growth only under an 8 h photoperiod. The µ~max~ was `r PURPARFits %>% filter(Strain == "SS120", WL == "450", O2_uM == "25", Photoperiod == 8) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ (Table. \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons m^-2^ s^-1^ blue light and 8 h photoperiod. The 25 µM O~2~, less than 16 h photoperiod and 180 µmol photons m^-2^ s^-1^  under white LED light experiments were not performed due to time constraints.

SS120 did not reproducibly grow when sparged to the lowest O~2~ of 2.5 µM (Figure \@ref(fig:SS120PAR)). 2.5 µM O~2~ growth experiments were not conducted for 4 and 16 h photoperiods under PAR of 180 µmol photons m^-2^ s^-1^, as no growth occurred when SS120 was exposed to 8 and 12 h photoperiods. Red light 16 h photoperiod experiments were not performed due to time constraints.
\newpage

@@

The GAM model in Figure \@ref(fig:SS120GAM) summarizes growth responses of SS120 to red (A) or blue (B) peak PAR and photoperiod, across the 2 imposed oxygen concentrations. Under 250 µM O~2~, Figure \@ref(fig:SS120GAM)A showed highest growth rates below PAR of 50 µmol photons m^-2^ s^-1^ and photoperiods between 8 and 12 h, indicated by the contour line labeled 0.29 d^-1^ (representing the 90^th^ percentile of achieved growth rate).  Under 250 µM O~2~ SS120 is constrained to deeper ocean waters through its intolerance of higher blue PAR levels. These findings align with @mooreComparativePhysiologySynechococcus1995 and are expected for a low light clade.  The disjunct regions of the GAM plot results from variable growth success of SS120 under 250 µM O~2~.  Growth rate patterns under red light were similar, although somewhat faster. In contrast, under 25 µM O~2~ and a photoperiod of 8 h SS120 exploited all blue peak PAR levels, achieving faster growth rates at a higher PAR of \~100 µmol photons m^-2^ s^-1^, indicated by the contour line labeled 0.5 d^-1^ (representing the 90^th^ percentile of achieved growth rate), out pacing the 90^th^ percentile fastest growth rates under 250 µM O~2~. Under red light and 25 µM O~2~ light SS120 grew across most conditions of peak PAR and photoperiod, achieving fastest growth under long photoperiods and peak PAR below 50 \~100 µmol photons m^-2^ s^-1^.  Thus, the designation of SS120 as a LL strain is dependent upon the [O~2]. SS120 did not, however, grow reliably under tested conditions at 2.5 µM O~2~.

@@

Under 250 µM O~2~ growth rates of *P. marinus* MIT9313, Clade LLIV, increased with longer
photoperiods, under low 30 µmol photons m^-2^ s^-1^ peak PAR, (Figure \@ref(fig:MIT9313PAR)).  Under intermediate 90 µmol photons m^-2^ s^-1^ peak PAR growth rates decreased with increasing blue light photoperiod. Blue light did not induce growth at 180 µmol photons m^-2^ s^-1^ peak PAR, while MIT9313 showed only marginal growth under white LED and red light at 180 µmol photons m^-2^ s^-1^ peak PAR, under the 8 h photoperiod, consistent with [@moorePhotophysiologyMarineCyanobacterium1999]. The µ~max~ was `r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under 30 µmol photons m^-2^ s^-1^ blue light and 16 h photoperiod.

For MIT9313 under 25 µM O~2~ growth rate increased with increasing photoperiods for all spectral wavebands tested (Figure \@ref(fig:MIT9313PAR)), with the fastest overall growth rate for MIT9313 
`r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "25", Photoperiod == 16, Par_ue == 90) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under peak PAR of 90 µmol photons m^-2^ s^-1^ and 16 h blue light photoperiod. In marked contrast to the 250 µM O~2~ growth experiments, MIT9313 grew when exposed to peak PAR of 180 µmol photons m^-2^ s^-1^ and blue light under all photoperiods except 16 h; additionally, white LED and red light treatments induced growth across all tested photoperiods under 25 µM O~2~. The 25 µM O~2~, 4 h photoperiod experiments under white LED light and were not performed due to time constraints.

MIT9313 grew under 2.5 µM O~2~  particularly under blue LED light, albeit generally slower than under the parallel experiments at 25 µM O~2~  (Figure \@ref(fig:MIT9313PAR)). Growth estimates showed scatter among replicates, suggesting  2.5 µM O~2~  is near the tolerance limit for growth of MIT9313. Growth rates increased with longer photoperiods under blue light treatments under peak PAR of 90 µmol photons m^-2^ s^-1^ . Growth  for MIT9313 under PAR of 180 µmol photons m^-2^ s^-1^ and blue light treatment decreased with increasing photoperiods with full growth inhibition under a 16 h photoperiod. The red light peak PAR of 180 µmol photons m^-2^ s^-1^ showed similar growth rates for 8 and 12 h photoperiods. The µ~max~ was `r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "2.5", Photoperiod == 4, Par_ue == 180) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under 4 h blue light photoperiod and PAR of 180 µmol photons m^-2^ s^-1^.  XXXXTHIS DOES NOT LOOK correct; 12 h, 90 uE photons blue light looks higher. The 2.5 µM O~2~ white LED treatments under 4, 8 and 16 h photoperiods and red light under 4 and 16 h photoperiods were not performed due to time constraints.

@@

@@

Cumulative diel PUR can potentially collapse photoperiod, PAR and spectral wavebands to a common metric of photosynthetically active light absorbed per day. Cumulative diel PUR dose (µmol photons m^-2^ d^-1^) was calculated from the imposed PUR (µmol photons m^-2^ s^-1^) and photoperiod (h). We plotted growth rates vs. cumulative diel PUR to determine whether growth is a simple response to diel PUR, across imposed photoperiods and spectral wavebands, or whether spectral wavebands or photoperiods have specific or interactive influences on growth beyond cumulative diel PUR.

Due to the absorption of *P. marinus* pigments in the blue spectral waveband range, the maximum imposed cumulative diel PUR under blue light is almost 3 times that of white LED light, and about 6 times the red light treatment, despite being derived from the same imposed photoperiods and peak PAR regimes. As such, only blue light experiments extend beyond cumulative diel PUR of ~ 2 x 10^6^ µmol photons m^-2^ d^-1^. This bias in the range of data leads to caution in comparing model fits of growth in response cumulative diel PUR under red, vs. blue, wavebands.

@@

The representative of HLI clade, *P. marinus* MED4, showed no growth under any 4 h photoperiod treatments, even when a 4 h photoperiod delivered cumulative diel PUR equivalent to other photoperiod treatments. In parallel MED4 showed no growth under 2.5 µM O~2~, no matter the level of diel cumulative PUR. 
In contrast, under 250 or 25  µM O~2~, and under any photoperiod greater than 4 h, MED4 growth under blue light was well described by a saturating response (REFERENCE for EQUATION) of growth to increasing cumulative diel blue PUR, with saturation of growth rate achieved around  1.0 x 10^6^ µmol m^-2^d^-1^ (Figure \@ref(fig:AllStrainPurFitsPlots)A), and no evidence of inhibition of growth at any achieved cumulative diel PUR.  Under the 'artificial' growth treatment of red light, MED4 achieved more growth per unit diel cumulative PUR  \@ref(fig:AllStrainPurFitsPlots)A), consistent with [@murphyPhotoinactivationPhotosystemII2017], who showed a lower cost for growth under red light, for MED4, because red light provokes less photoinactivation of PSII, than equivalent levels of blue light.  XXXXDISTINCT PHOTOPERIOD FITS REFER to SupplementalXXXXX

@@

The representative of the LLIII clade, *P. marinus* SS120 showed almost no growth under 2.5 µM O~2~ experiments. Most 4 h photoperiod treatments of SS120 also did not grow under 250 µM O~2~, even when a 4 h photoperiod delivered cumulative diel PUR equivalent to other photoperiod treatments. SS120 did not grow when exposed to more than ~1.0 x 10^6^ µmol photons m^-2^ d^-1^ of cumulative diel PUR under any spectral waveband or photoperiod combination, under 250 µM O~2~ experiments(Figure \@ref(fig:AllStrainPurFitsPlots)B).

Under both 25 or 250 µM O~2~ experiments, SS120 growth plateaued by about 5.0 x 10^5^ µmol photons m^-2^ d^-1^ diel PUR, with some scatter among photoperiod and spectral waveband regimes. The onset of growth inhibition extended to higher cumulative diel PUR for cultures under 25 µM O~2~, showing that SS120 is partially protected from photoinhibition of growth by 25 µM O~2~.  Under 25 µM O~2~ red light again generated more growth of SS120 per unit cumulative diel PUR, than did blue light, again consistent with lower cost of growth through lower photoinactivation under red light.
XXXXDISTINCT PHOTOPERIOD FITSXXXXX

@@

The LLIV clade representative, *P. marinus* MIT9313, showed growth rising to a plateau by about 5 x 10^5^ µmol photons m^-2^ d^-1^ of cumulative diel PUR, with higher growth rates over a narrower plateau under 25 and 250 µM O~2~, compared to a wider, lower, flatter response to cumulative diel PUR under 2.5 µM O~2~. Above about 1.0 x 10^6^ µmol photons m^-2^ d^-1^ of cumulative PUR under 250 µM O~2~, MIT9313 showed full inhibition of growth, across photoperiods, and spectral wavebands.  In contrast, under 25 µM O~2~, MIT9313 showed a greatly extended exploitation of higher cumulative diel PUR, with full growth inhibition only above about 3.5 x 10^6^ µmol photons m^-2^ d^-1^.  Similarly, under 2.5 µM O~2~, MIT9313 grew more slowly, but only showed full growth inhibition above about 3.5 x 10^6^ µmol photons m^-2^ d^-1^ cumulative diel PUR (Figure \@ref(fig:AllStrainPurFitsPlots)C).

Our data support enhanced growth under conditions of low cumulative dielPUR and 660 nm (red) spectral bandwidth, as supported by [@murphyPhotoinactivationPhotosystemII2017] who found a lower cost of growth, due to decreased photoinactivation of PSII under red, compared to blue wavebands.  Interestingly, this protective effect of red light disappears for MIT9313 growing under 2.5 µM O~2~, possibly because photoinactivation is strongly suppressed under this low [O~2~].
XXXXDISTINCT PHOTOPERIOD FITSXXXXX
